\section{Theoretical Analysis}
\label{sec:theory}

We provide a formal interpretation of the self-conditioning dynamics observed in Section~\ref{sec:method}, explaining why rotation converges while translation diverges, and identifying the mechanistic pathway through which conditioning operates.

\subsection{Contraction Mapping on SO(3)}

Let $f: \text{SO}(3)^N \rightarrow \text{SO}(3)^N$ denote the composite mapping from conditioning rotations to Pi3X's output rotations after pipeline processing (gauge fix, outlier rejection, Fréchet mean). We define the geodesic distance on SO(3) as:
\begin{equation}
    d_{\text{SO}(3)}(R_1, R_2) = \|\log(R_1^\top R_2)\|_F
\end{equation}
where $\log: \text{SO}(3) \rightarrow \mathfrak{so}(3)$ is the matrix logarithm. This equals the angle of the relative rotation $R_1^\top R_2$.

If $f$ is a contraction mapping, i.e., there exists $\lambda < 1$ such that for all $R$ in a neighbourhood of the true rotations $R^*$:
\begin{equation}
    d_{\text{SO}(3)}(f(R), R^*) \leq \lambda \cdot d_{\text{SO}(3)}(R, R^*)
\end{equation}
then by the Banach fixed-point theorem, the iteration $R^{(n+1)} = f(R^{(n)})$ converges to a unique fixed point $R^\infty$ satisfying $f(R^\infty) = R^\infty$.

\paragraph{Empirical measurement.} We measure the contraction rate by feeding GT poses corrupted with known noise levels and comparing input error to output error (Section~\ref{sec:ablation}, Table~\ref{tab:noise_robustness}):
\begin{itemize}
    \item Input error 4.4\degree{} $\rightarrow$ output error 3.5\degree{} ($\lambda \approx 0.80$)
    \item Input error 5.0\degree{} $\rightarrow$ output error 3.3\degree{} ($\lambda \approx 0.67$)
    \item Input error 2.0\degree{} $\rightarrow$ output error 1.6\degree{} ($\lambda \approx 0.80$)
\end{itemize}

The contraction rate is approximately $\lambda \approx 0.8$ in the 2--5\degree{} regime. This predicts convergence to a fixed point at:
\begin{equation}
    \text{ARE}^* = \frac{\lambda}{1-\lambda} \cdot \epsilon_{\text{irr}} \approx 4 \cdot \epsilon_{\text{irr}}
    \label{eq:fixed_point}
\end{equation}
where $\epsilon_{\text{irr}}$ is the model's irreducible error (the residual even with perfect conditioning). With $\text{ARE}^* \approx 3.1\degree{}$ observed experimentally, this gives $\epsilon_{\text{irr}} \approx 0.8\degree{}$, consistent with the 1.2\degree{} achieved under perfect GT conditioning (which includes additional pipeline processing noise).

\paragraph{Basin of attraction.} The contraction property holds only within a finite neighbourhood of $R^*$. Our noise robustness experiments (Section~\ref{sec:ablation}) show that conditioning with $>10\degree{}$ error causes catastrophic failure (output error jumps to 180\degree{}). This defines a basin of attraction of approximately $\pm 10\degree{}$ around the true solution---our initial estimate (4.4\degree{} error) lies safely within it.

\subsection{Failure of Contraction for Translation}

Pi3X outputs poses in Sim(3), not SE(3). The translation component $\mathbf{t}^c$ is meaningful only relative to the per-run scale $s$. When fed back as SE(3) conditioning, a scale mismatch accumulates:
\begin{equation}
    \mathbf{t}^{(n+1)} = s^{(n)} \cdot g(\mathbf{t}^{(n)}, R^{(n)}) + \boldsymbol{\beta}
\end{equation}
where $s^{(n)} \neq 1$ in general and $\boldsymbol{\beta}$ represents systematic bias. Since the scale factor $s^{(n)}$ varies per iteration and is not corrected, this does not satisfy contraction:
\begin{equation}
    \|\mathbf{t}^{(n+1)} - \mathbf{t}^*\| \not\leq \lambda \|\mathbf{t}^{(n)} - \mathbf{t}^*\|
\end{equation}

\paragraph{Mechanistic explanation.} The encode path uses conditioning translations to normalise depth scale: $\mathbf{t}' = \mathbf{t} / \text{median}(\|\mathbf{t}\|)$ before constructing ray embeddings. Since Pi3X outputs in Sim(3), the scale of the output translation is arbitrary relative to the conditioning scale. Feeding output translations back as input creates a compounding scale mismatch---each iteration's normalisation references a different implicit scale, causing monotonic drift. Rotation is immune because it is scale-invariant: the rotation subgroup of Sim(3) equals that of SE(3).

This analysis motivates our hybrid fusion strategy: rotation contracts and should be iterated; translation does not and should be anchored to the unconditioned estimate.

\subsection{Block Coordinate Descent Interpretation}
\label{sec:bcd}

Self-conditioning admits a natural interpretation as block coordinate descent (BCD)~\cite{wright2015} on an implicit joint objective over poses and features. Let $\mathcal{F}(R, \phi)$ denote the implicit objective that Pi3X minimises during pretraining, where $R \in \text{SO}(3)^N$ are camera rotations and $\phi$ are internal feature representations. Each iteration alternates between two steps:
\begin{enumerate}
    \item \textbf{Feature update} (given poses): The encode path constructs ray embeddings $\mathbf{r}_{i} = \text{lift}(\boldsymbol{\Pi}^c) \cdot [R^{(n)} \mid \mathbf{t}]$ and normalises depth scale using $\|\mathbf{t}\|$, producing improved features $\phi^{(n+1)} = h(R^{(n)}, \text{imgs})$.
    \item \textbf{Pose update} (given features): The decoder estimates poses from the enriched features: $R^{(n+1)} = g(\phi^{(n+1)})$.
\end{enumerate}

The composite map $f = g \circ h$ maps conditioning poses to output poses. Contraction ($\lambda < 1$) arises because ray embeddings provide angular priors that collapse directional ambiguity in the feature space---each iteration's features are strictly more informative than the previous iteration's, yielding strictly better pose estimates until the fixed point.

\paragraph{Why rotation contracts but translation diverges.} Ray embeddings encode pixel-to-3D \emph{direction}, which depends on rotation but is scale-invariant. Better rotation priors yield better directional features, which yield better rotation estimates---a self-reinforcing loop with diminishing returns (contraction). Translation enters only through scale normalisation, where the Sim(3) ambiguity prevents convergence (Section~4.2). This explains the $\lambda \approx 0.8$ contraction rate as the composite Lipschitz constant of the encode-decode loop restricted to SO(3).

\paragraph{Consistency with ablation evidence.} The BCD interpretation predicts that the active mechanism is the encode path (where poses $\rightarrow$ features), not the decode-path PoseInjectBlocks (where poses modulate attention). Our layer ablation (Section~\ref{sec:ablation_mechanism}) confirms this: zeroing all five PoseInjectBlocks produces identical accuracy to leaving them active (ARE 3.84\degree{} vs 3.85\degree{}), while removing encode-path conditioning entirely (via identity poses) eliminates all improvement. The decode-path blocks are vestigial for this task.

\subsection{Posterior Unimodality Under Self-Conditioning}

One might expect that perturbing the conditioning signal and averaging---Monte Carlo integration over the model's implicit posterior $p(R^* \mid R + \boldsymbol{\epsilon}, \text{imgs})$---would provide additional variance reduction. If the posterior has non-trivial spread, averaging $S$ samples should reduce error as $O(1/\sqrt{S})$.

However, the BCD interpretation explains why this fails: since the mechanism is deterministic feature enrichment (ray embeddings from poses), not stochastic sampling, all perturbations within the basin of attraction produce the same enriched features and converge to the same output. The posterior is effectively a point mass (angular standard deviation $<$0.02\degree{} across samples), leaving no variance for MC averaging to reduce. Section~\ref{sec:ablation_S} confirms empirically that $S=1$ is optimal.
