\section{Foundation Geometry Model Benchmark}
\label{sec:benchmark}

To select the best backbone for our calibration pipeline, we benchmark seven recent foundation geometry models on our dataset. All models receive the same 8 synchronised camera images per timestep and must output camera poses without any prior information.

\subsection{Experimental Protocol}

We sample 91 timesteps at regular intervals from synchronised video (2688$\times$1520 resolution, resized per model requirements). For each model, we:
\begin{enumerate}
    \item Run inference on all timesteps (8 cameras per timestep)
    \item Aggregate per-timestep poses via quaternion averaging (rotations) and median (translations)
    \item Apply Sim(3) Procrustes alignment to ground truth (with reflection check)
    \item Compute ATE, ARE, RPE, and co-located pair distances
\end{enumerate}

Ground truth poses are obtained from COLMAP applied to a static scene capture (no cattle present). Two camera pairs (C2--C3 and C5--C6) share physical mounting poles, providing a structural validation metric (ground truth distance $\approx$ 0).

\subsection{Results}

\begin{table}[t]
\centering
\caption{Foundation geometry model benchmark on our wide-baseline, feature-adversarial dataset. Models are grouped by type (N-view joint vs.\ pairwise with post-hoc alignment) and ranked by ATE within each group. ``Cond.'' indicates whether the model accepts pose/intrinsic priors as conditioning input---a prerequisite for our self-refinement method. All metrics use Sim(3)-aligned predictions; C2C3 measures co-located camera consistency. Pairwise models use only 20 timesteps due to computational cost.}
\label{tab:benchmark}
\begin{tabular}{@{}llcccccccc@{}}
\toprule
Model & Venue & Type & Cond. & ATE$\downarrow$ & ARE$\downarrow$ & RPE$_R$$\downarrow$ & RPE$_t$$\downarrow$ & C2C3$\downarrow$ & Time (s) \\
\midrule
\multicolumn{10}{@{}l}{\textit{N-view (joint)}} \\[2pt]
\textbf{Pi3X} & CVPR'25 & N-view & \checkmark & \textbf{0.177} & 4.37\degree & \textbf{6.26\degree} & {0.39} & {0.088} & 106 \\
Pi3 & CVPR'25 & N-view & --- & 0.242 & \textbf{4.08\degree} & 5.98\degree & \textbf{0.37} & \textbf{0.000} & 121 \\
MapAnything & CVPR'25 & N-view & \checkmark & 0.627 & 9.17\degree & 14.19\degree & 0.93 & 0.313 & 86 \\
VGGT & CVPR'25 & N-view & --- & 1.248 & 10.48\degree & 12.82\degree & 1.55 & 2.011 & 75 \\
NoPo4D & arXiv'25 & N-view & --- & 2.556 & 62.06\degree & 77.01\degree & 3.17 & 0.627 & 214 \\
Fast3R & CVPR'25 & N-view & --- & 2.743 & 96.64\degree & 102.88\degree & 4.36 & 3.183 & 460 \\
\midrule
\multicolumn{10}{@{}l}{\textit{Pairwise (post-hoc alignment)}} \\[2pt]
Pow3R & CVPR'25 & Pairwise & \checkmark & 1.835 & 180\degree & 28.53\degree & 7.24 & 0.218 & 497 \\
MonST3R & ICLR'25 & Pairwise & --- & 2.607 & 47.24\degree & 23.87\degree & 3.56 & 0.111 & 1492 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Analysis}

The results reveal a clear performance gap between architectures. Pi3X and Pi3 (its smaller variant) both achieve sub-5\degree{} rotation error, with Pi3 achieving slightly better ARE (4.08\degree{} vs.\ 4.37\degree{}) but worse ATE (0.242 vs.\ 0.177). MapAnything and VGGT form a second cluster at $\sim$10\degree{} rotation error: functional but substantially less accurate. The remaining models (NoPo4D, Fast3R, MonST3R, Pow3R) all exceed 20\degree{} rotation error and effectively fail on this dataset.

Two structural patterns emerge from these results. First, N-view joint architectures (Pi3X, MapAnything, VGGT) consistently outperform pairwise models, confirming that joint multi-view reasoning is essential when inter-camera overlap is minimal---pairwise models cannot propagate geometric constraints across views that share no visual content. Second, the co-location metric reveals differences in structural understanding that are masked by aggregate error: Pi3X correctly infers that cameras C2 and C3 share a mounting pole (distance 0.088), while VGGT (2.011) and Fast3R (3.183) place them far apart.

Of the eight models tested, only three support pose conditioning (Table~\ref{tab:benchmark}, ``Cond.'' column): Pi3X, MapAnything, and Pow3R. Pi3 (base), despite competitive rotation accuracy, lacks a conditioning interface and thus cannot benefit from self-refinement. Pi3X combines the best overall accuracy with conditioning support, making it the natural backbone. Note that Pi3 (base) does not support conditioning despite being from the same model family---the conditioning mechanism was introduced in the Pi3X extension. The remainder of this paper develops methods to push beyond Pi3X's $\sim$4.4\degree{} accuracy ceiling, which resists further improvement through standard geometric post-processing alone.
