\section{Ablation Studies}
\label{sec:ablation}

We isolate the contribution of each design decision through four ablation experiments.

\subsection{Effect of Hybrid Decoupling}

Table~\ref{tab:naive_vs_hybrid} compares na\"ive iterative refinement (feeding the full conditioned output back without decoupling) against our hybrid approach.

\begin{table}[h]
\centering
\caption{Na\"ive vs.\ hybrid iterative refinement. Na\"ive iteration improves rotation at the cost of translation; hybrid fusion achieves superior rotation improvement while preserving translation accuracy.}
\label{tab:naive_vs_hybrid}
\begin{tabular}{@{}lcccc@{}}
\toprule
Method & ATE$\downarrow$ & ARE$\downarrow$ & $\Delta$ATE & $\Delta$ARE \\
\midrule
Unconditioned baseline & 0.161 & 4.39\degree{} & --- & --- \\
Na\"ive self-refine (iter 5) & 0.195 & 3.60\degree{} & +21\% & $-$18\% \\
Hybrid MC (iter 2) & 0.160 & 3.13\degree{} & $-$1\% & $-$29\% \\
\bottomrule
\end{tabular}
\end{table}

Na\"ive iteration degrades ATE by 21\% while improving ARE by only 18\%. In contrast, hybrid fusion achieves 29\% ARE improvement with negligible translation change. This confirms the theoretical prediction of Section~\ref{sec:theory}: translation does not contract under self-conditioning and must be anchored to the unconditioned estimate.

\subsection{Sensitivity to Conditioning Quality}
\label{sec:noise_sensitivity}

To characterise the basin of attraction, we condition Pi3X with ground truth poses corrupted by varying levels of Gaussian noise (Table~\ref{tab:noise_robustness}).

\begin{table}[h]
\centering
\caption{Effect of conditioning pose quality on output accuracy. Translation noise is isotropic Gaussian in scene coordinates. Beyond $\sim$10\degree{} rotation noise, conditioning becomes counterproductive and the model diverges.}
\label{tab:noise_robustness}
\begin{tabular}{@{}cccc@{}}
\toprule
Rot.\ noise (\degree{}) & Trans.\ noise & ATE$\downarrow$ & ARE$\downarrow$ \\
\midrule
0 (perfect GT) & 0 & 0.097 & 1.22\degree{} \\
1 & 0.05 & 0.087 & 1.30\degree{} \\
2 & 0.10 & 0.080 & 1.60\degree{} \\
5 & 0.20 & 0.077 & 3.34\degree{} \\
10 & 0.50 & 0.368 & 180\degree{} \\
20 & 1.00 & 0.951 & 180\degree{} \\
\bottomrule
\end{tabular}
\end{table}

The model tolerates conditioning noise up to approximately 5\degree{} of rotation error, producing output that improves upon the conditioning input. At 10\degree{}, the conditioning signal becomes misleading and the model collapses to degenerate solutions ($180\degree{}$ ARE indicates random orientations). The transition is sharp rather than gradual, indicating a well-defined basin boundary. Our self-conditioned prior ($\sim$4.4\degree{} error) lies near the edge of this basin, which explains both why self-refinement succeeds and why it saturates after 2--3 iterations as the estimate approaches the basin's contraction limit.

\subsection{Convergence Behaviour}

Table~\ref{tab:convergence} reports per-iteration metrics for the full hybrid MC method.

\begin{table}[h]
\centering
\caption{Convergence of iterated hybrid MC refinement ($S=4$ samples per iteration). Iteration~2 achieves the minimum ARE; iteration~4 produces an orientation flip in the rotation estimate.}
\label{tab:convergence}
\begin{tabular}{@{}ccccc@{}}
\toprule
Iteration & ATE & ARE (\degree{}) & RPE$_R$ (\degree{}) & RPE$_t$ \\
\midrule
0 (unconditioned) & 0.161 & 4.39 & 6.26 & 0.375 \\
1 & 0.162 & 3.43 & 4.88 & 0.321 \\
2 & 0.160 & 3.13 & 4.58 & 0.309 \\
3 & 0.163 & 3.38 & 4.70 & 0.307 \\
4 & 0.160 & 180\degree{} & 4.78 & 8.55 \\
\bottomrule
\end{tabular}
\end{table}

The method converges monotonically through iterations 1--2, achieving a 29\% ARE reduction. Iteration~3 shows slight regression, and iteration~4 produces catastrophic failure (orientation flip in at least one camera). This behaviour is consistent with the basin-of-attraction analysis: as the conditioned estimate approaches the $\sim$3\degree{} fixed point, the margin between the estimate's error and the basin boundary narrows, making the iteration sensitive to perturbation noise. We adopt a conservative stopping criterion of $M=2$ iterations or ARE improvement below $0.1\degree{}$ between successive iterations.

\subsection{Inefficacy of Sim(3) Re-alignment}

One might hypothesise that translation degradation under na\"ive iteration is caused by scale drift---each iteration's output uses a different arbitrary scale, and this accumulates. To test this, we Sim(3)-align each iteration's output back to the conditioning frame before re-injection. Table~\ref{tab:sim3} shows that this intervention has no measurable effect.

\begin{table}[h]
\centering
\caption{Sim(3)-constrained vs.\ unconstrained na\"ive iteration. Constraining scale does not prevent translation degradation, confirming that the failure mechanism is systematic conditioning bias rather than scale drift.}
\label{tab:sim3}
\begin{tabular}{@{}ccccc@{}}
\toprule
Iter & Sim(3) ATE & Sim(3) ARE & Unconstrained ATE & Unconstrained ARE \\
\midrule
0 & 0.152 & 4.69\degree{} & 0.152 & 4.69\degree{} \\
1 & 0.160 & 3.50\degree{} & 0.160 & 3.50\degree{} \\
2 & 0.180 & 3.31\degree{} & 0.179 & 3.29\degree{} \\
5 & 0.197 & 3.61\degree{} & 0.195 & 3.59\degree{} \\
\bottomrule
\end{tabular}
\end{table}

The near-identical degradation trajectories confirm that translation failure is not a scale problem but a structural one: Pi3X's conditioning mechanism introduces systematic translational bias that is independent of the output's global scale. This motivates our decision to anchor translations entirely to the unconditioned source rather than attempting to correct the conditioned output.

\subsection{Effect of Sample Count $S$}
\label{sec:ablation_S}

We fix $M=2$ iterations and $\sigma_R = 2\degree{}$ and vary the number of MC samples $S$ per iteration (Table~\ref{tab:s_ablation}).

\begin{table}[h]
\centering
\caption{Effect of MC sample count $S$ on final accuracy ($M=2$, $\sigma_R=2\degree{}$). The angular standard deviation across samples measures the spread of the per-sample rotation estimates. $S=1$ (deterministic, no perturbation) achieves the lowest ARE.}
\label{tab:s_ablation}
\begin{tabular}{@{}cccccc@{}}
\toprule
$S$ & ATE$\downarrow$ & ARE (\degree{})$\downarrow$ & Ang.\ Std (\degree{}) & Cost & Time (s) \\
\midrule
1 & \textbf{0.151} & \textbf{3.19} & --- & 3$\times$ & 214 \\
2 & 0.154 & 3.43 & $<$0.01 & 5$\times$ & 429 \\
4 & 0.156 & 3.24 & 0.38 & 9$\times$ & 899 \\
8 & 0.154 & 3.48 & 0.55 & 17$\times$ & 1714 \\
16 & 0.152 & 3.41 & 0.58 & 33$\times$ & 3430 \\
\bottomrule
\end{tabular}
\end{table}

Counter to the theoretical prediction that averaging $S$ samples should reduce variance as $O(1/\sqrt{S})$, increasing $S$ provides no improvement---and in most cases slightly degrades---ARE relative to $S=1$. The angular standard deviation across MC samples is very small ($<$1\degree{} for all $S$), confirming that the model's posterior under our conditioning regime is tightly concentrated around a single mode. In this regime, stochastic perturbation adds noise to the conditioning signal without exploring meaningfully different solutions, and the Fréchet mean over perturbed outputs is slightly less accurate than the single deterministic conditioned pass.

This finding revises our earlier recommendation: for this dataset, the optimal configuration is $S=1$ (no perturbation), $M=2$ iterations, yielding ARE~=~3.19\degree{} at only 3$\times$ the cost of a single unconditioned pass---a 27\% rotation improvement with minimal compute overhead.

\subsection{Weighting Strategy for Sample Aggregation}
\label{sec:ablation_weighting}

Given that MC sampling provides marginal benefit, we test whether non-uniform weighting can extract better estimates from multi-sample runs ($S=8$, $M=1$). Table~\ref{tab:weighting} compares four weighting strategies.

\begin{table}[h]
\centering
\caption{Weighting strategies for MC sample aggregation ($S=8$, $M=1$). All strategies produce near-identical results, confirming that per-sample differences are negligible.}
\label{tab:weighting}
\begin{tabular}{@{}lcccc@{}}
\toprule
Strategy & ATE$\downarrow$ & ARE (\degree{})$\downarrow$ & RPE$_R$ (\degree{})$\downarrow$ & RPE$_t$$\downarrow$ \\
\midrule
Equal weight (baseline) & 0.152 & 4.06 & 5.06 & 0.398 \\
Confidence-weighted & 0.152 & 4.06 & 5.07 & 0.398 \\
Inverse-variance & 0.152 & 4.04 & 5.05 & 0.397 \\
Robust reweighted & 0.152 & 4.04 & 5.06 & 0.397 \\
\bottomrule
\end{tabular}
\end{table}

All four strategies produce functionally identical results (ARE differences $<$0.02\degree{}). This confirms the finding from Section~\ref{sec:ablation_S}: the posterior is unimodal and tightly concentrated, so all samples carry essentially equal information regardless of their Pi3X confidence score or internal consistency. The equal-weight assumption is not a limiting approximation in this regime---the limitation is that multiple samples provide no additional information beyond a single pass.
