\section{Results}
\label{sec:results}

\subsection{Experimental Setup}

\paragraph{Dataset.} We evaluate on 8 static cameras (2688$\times$1520 resolution) deployed in a cattle feedlot. We sample $N_t = 91$ timesteps at regular intervals from synchronised video. Ground truth poses are obtained from COLMAP applied to a static capture of the same scene (without cattle).

\paragraph{Evaluation metrics.} All methods are evaluated after Sim(3) Procrustes alignment to ground truth (with reflection check):
\begin{itemize}
    \item \textbf{ATE}: Absolute Translation Error (RMSE of camera position errors)
    \item \textbf{ARE}: Absolute Rotation Error (mean geodesic distance to GT rotations, in degrees)
    \item \textbf{RPE rot / trans}: Relative Pose Error averaged over all $\binom{8}{2} = 28$ camera pairs
    \item \textbf{C2C3}: Distance between co-located cameras C2 and C3 (GT $\approx$ 0; structural consistency test)
\end{itemize}

Note that ATE and RPE trans are in arbitrary units (scene scale is determined by the Sim(3) alignment, not metric ground truth).

\subsection{Comprehensive Comparison}

Table~\ref{tab:comparison} presents the progressive improvement from raw Pi3X output through our full method.

\begin{table}[t]
\centering
\caption{Progressive improvement from raw Pi3X through our full method. Cost is relative to a single unconditioned inference pass ($\sim$100s on an A100). Row~3 shows na\"ive self-conditioning (rotation improves but translation degrades). Row~4 introduces hybrid fusion (translation preserved). Row~5 iterates the hybrid method twice ($M$=2).}
\label{tab:comparison}
\begin{tabular}{@{}lcccccc@{}}
\toprule
Method & ATE$\downarrow$ & ARE$\downarrow$ & RPE$_R$$\downarrow$ & RPE$_t$$\downarrow$ & C2C3$\downarrow$ & Cost \\
\midrule
1. Raw Pi3X (quat average) & 0.177 & 4.37\degree & 6.26\degree & 0.395 & 0.088 & 1$\times$ \\
2. + Pipeline (Stage 1) & 0.161 & 4.39\degree & 6.26\degree & 0.375 & 0.000 & 1$\times$ \\
3. + Self-Conditioning (Stage 2) & 0.166 & 3.49\degree & 4.93\degree & 0.384 & 0.000 & 2$\times$ \\
4. + Hybrid Fusion (Stage 3) & 0.159 & 3.45\degree & 4.93\degree & 0.322 & 0.000 & 2$\times$ \\
\textbf{5. + Iterated Hybrid ($M$=2)} & \textbf{0.151} & \textbf{3.19\degree} & \textbf{3.84\degree} & \textbf{0.254} & \textbf{0.000} & 3$\times$ \\
\midrule
\rowcolor{gray!15}
\textit{Oracle (GT pose conditioning)} & \textit{0.097} & \textit{1.22\degree} & \textit{---} & \textit{---} & \textit{---} & 1$\times$ \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Analysis}

The progressive results in Table~\ref{tab:comparison} reveal several findings.

\paragraph{Rotation error is not reducible by classical aggregation.} The pipeline (row~2) reduces ATE by 10\% through pose graph optimization with co-location constraints, but provides no rotation improvement ($4.37\degree \rightarrow 4.39\degree$). This confirms that Pi3X's rotation error is systematic model bias rather than statistical noise: robust averaging over 91 timesteps has already extracted all available information from the raw outputs.

\paragraph{Self-conditioning exhibits rotation--translation asymmetry.} Conditioning on the pipeline estimate (row~3) reduces ARE by 20\% but increases ATE by 3\%. This asymmetry---the central observation motivating our method---is consistent across multiple runs and noise configurations (Section~\ref{sec:ablation}).

\paragraph{Hybrid fusion resolves the asymmetry.} Decoupling rotation and translation channels (row~4) achieves $3.45\degree{}$ ARE (capturing the conditioned rotation improvement) while maintaining ATE at 0.159 (preserving the unconditioned translation quality). The computational overhead is minimal: one additional conditioned inference pass ($2\times$ total cost).

\paragraph{Iteration further improves rotation at minimal cost.} The full method (row~5) applies hybrid fusion twice ($M=2$), reducing ARE by an additional 8\% over single-step hybrid fusion ($3.45\degree \rightarrow 3.19\degree$). Our ablation study (Section~\ref{sec:ablation_S}) further investigates whether Monte Carlo perturbation of the conditioning signal can provide additional gains; it does not, revealing that the model's posterior is tightly unimodal. The $3\times$ total cost makes iterated hybrid fusion practical even for interactive applications.

\paragraph{Aggregate improvement.} Comparing the raw model output (row~1) to the full method (row~5): ARE improves by 27\% ($4.37\degree \rightarrow 3.19\degree$), RPE rotation by 39\%, RPE translation by 36\%, and ATE by 15\%. The co-location constraint (C2C3) is satisfied exactly under pose graph optimization (rows~2--5).

\paragraph{Oracle upper bound.} The final row reports Pi3X conditioned on ground truth poses, representing the best achievable accuracy given perfect conditioning. This oracle achieves ARE\,=\,1.22\degree{} and ATE\,=\,0.097---reductions of 72\% and 40\% over the unconditioned baseline. The gap between our self-refined result (3.19\degree{}) and the oracle (1.22\degree{}) quantifies the remaining headroom attributable to conditioning quality. The monotonic improvement from unconditioned ($4.39\degree$) through self-conditioned ($3.19\degree$) to oracle ($1.22\degree$) confirms that the conditioning pathway provides genuine geometric information and that our method successfully extracts a substantial fraction of the available signal without external data.
