\section{Discussion}
\label{sec:discussion}

\paragraph{Practical deployment.} The iterated hybrid fusion method ($M=2$) provides the most favourable accuracy--cost tradeoff: two conditioned inference passes (3$\times$ total compute) yield 27\% rotation improvement while preserving translation. Monte Carlo perturbation of the conditioning signal adds compute without benefit---the model's posterior under self-conditioning is tightly unimodal (Section~\ref{sec:ablation_S}), making deterministic iteration strictly preferable. Hybrid decoupling is essential: na\"ive self-conditioning without it leads to progressive translation degradation.

\paragraph{Generality of the framework.} While our experiments use Pi3X, the decoupled hybrid refinement framework applies in principle to any learned geometry model that (i) accepts pose conditioning as input, and (ii) outputs poses defined up to a gauge ambiguity. These properties are shared by several recent architectures (Table~\ref{tab:benchmark}). The contraction analysis (Section~\ref{sec:theory}) provides a diagnostic: measuring input--output error reduction under corrupted conditioning reveals whether a given model exhibits the contraction property and, if so, at what rate. Models with $\lambda < 1$ are amenable to self-refinement; those with $\lambda \geq 1$ are not.

\paragraph{Limitations.} Several constraints bound the applicability of our method. The accuracy floor ($\sim$3\degree{} ARE) is determined by the contraction rate and the model's irreducible error---it cannot be lowered without stronger external priors or architectural modifications to the conditioning mechanism. Translation accuracy is bounded by the quality of unconditioned inference, since hybrid fusion anchors translations to this source. The convergence is empirical: while iterations 1--3 consistently improve, iteration 4 and beyond can produce orientation flips (Table~\ref{tab:convergence}), necessitating a stopping criterion. Finally, our evaluation is conducted on a single multi-camera installation; generalisation to other wide-baseline deployments with different camera counts and geometries remains to be validated.

\paragraph{Future work.} Several directions merit investigation. First, Pi3X also accepts depth conditioning---monocular depth priors (e.g., Depth Anything~v2) could provide complementary geometric signal without requiring ground truth poses, potentially breaking the rotation floor. Second, partial conditioning (conditioning only high-confidence cameras while leaving others unconditioned) may exploit the model's cross-view attention to propagate reliable pose information to uncertain views. Third, sparse geometric anchors (a pre-calibrated camera pair or LiDAR registration points) could provide a stronger initialisation that places the estimate deeper within the contraction basin, accelerating convergence and lowering the fixed point. Fourth, the tight unimodality of the posterior under self-conditioning suggests that the model has learned a near-deterministic mapping; investigating whether alternative perturbation strategies (e.g., perturbing image inputs rather than poses, or using dropout at inference) can elicit meaningful posterior diversity remains an open question.
